---
title: "ENEC891 Research Presentation"
author: "Grace Di Cecco"
date: "November 27, 2018"
output: github_document
---

```{r setup}
library(tidyverse)
library(sf)
library(raster)
library(SDMTools)
library(tmap)

# package foreach [for each region]
```

## Area of study

[Bird Conservation Regions](http://nabci-us.org/resources/bird-conservation-regions-map/) in continental US.

```{r}
#bcrshp <- read_sf("\\\\Bioark.bio.unc.edu\\hurlbertlab\\DiCecco\\bcr_terrestrial_shape\\BCR_Terrestrial_master.shp")
bcrshp <- read_sf("/Volumes/hurlbertlab/DiCecco/bcr_terrestrial_shape/BCR_Terrestrial_master.shp")
#BCRs

bcrs.us <- bcrshp %>%
  filter(!is.na(REGION), REGION == "USA", PROVINCE_S != "ALASKA", PROVINCE_S != "HAWAIIAN ISLANDS") %>%
  dplyr::select(BCR) %>%
  mutate(BCRfactor = as.factor(BCR))

map_bcrs <- tm_shape(bcrs.us) + tm_borders() + tm_fill(col = "BCRfactor", palette = "cat") + tmap_options(max.categories = 32)
map_bcrs

```

For my project, I looked at a subset of BCRs: 
9 - Great Basin
12 - Boreal Hardwood Transition
13 - Lower Great Lakes
14 - Atlantic Northern Forests
18 - Shortgrass Prairie
19 - Central Mixed Grass Prairie
23 - Prairie Hardwood Transition
27 - Southeastern Coastal Plain
29 - Piedmont

```{r}
bcrs <- c(9, 12, 13, 14, 18, 19, 23, 27, 29) # BCRs of interest
bcr_sub <- bcrs.us %>%
  filter(BCR %in% bcrs)

map_bcr_sub <- tm_shape(bcrs.us) + tm_borders()
map_bcr_sub + tm_shape(bcr_sub) + tm_fill(col = "BCRfactor", palette = "cat") + tm_borders() + tm_legend(show = F)

```

## Data source

NLCD dataset from 4 time points: 1992, 2001, 2006, 2011. Freely available for [download](http://nabci-us.org/resources/bird-conservation-regions-map/).

Divides land cover into [20 classes](https://www.mrlc.gov/data/legends/national-land-cover-database-2011-nlcd2011-legend). Some changes in classification between 1992 and 2000's: mainly in classification of developed land and shrubland. No changes in forest/grassland categories.

## Fragmentation measures

SDMtools package: enables calculation of Fragstats-like habitat fragmentation metrics from raster files in R.

```{r}
?ClassStat
?PatchStat
```

Analysis workflow:
- Merge NLCD rasters to continental US (when necessary)
- Crop NLCD raster with BCR shapefile
- Calculate fragmentation measures on each BCR region for each timepoint

## EDA/Plots of habitat fragmentation over time in study regions

```{r read in data}
#frags <- read.csv("\\\\Bioark.bio.unc.edu\\hurlbertlab\\DiCecco\\data\\fragmentation_indices_nlcd.csv", stringsAsFactors = F)
frags <- read.csv("/Volumes/hurlbertlab/DiCecco/data/fragmentation_indices_nlcd.csv", stringsAsFactors = F)
head(frags)
```

Matching land cover classes over time [https://pubs.usgs.gov/of/2008/1379/pdf/ofr2008-1379.pdf]

```{r}
classlegend00s <- data.frame(class = c(11:12, 21:24, 31, 41:43, 51, 52, 71, 81:82, 90, 95), 
                          legend = c("Open water", "Perennial ice/snow", "Developed, open space", "Developed, low intensity", "Developed, medium intensity", "Developed, high intensity", "Barren land", "Deciduous forest", "Evergreen forest", "Mixed forest", "Dwarf scrub", "Shrub/scrub", "Grassland/herbaceous", "Pasture/hay", "Cultivated crops", "Woody wetlands", "Emergent herbaceous wetlands"))
classlegend92 <- data.frame(class = c(11:12, 85, 21:23, 31:33, 41:43, 51, 61, 71, 81:84, 91:92),
                            legend = c("Open water", "Perennial ice/snow", "Developed, open space", "Developed, low intensity", "Developed, medium intensity", "Developed, high intensity", "Barren land", "Barren land", "Barren land", "Deciduous forest", "Evergreen forest", "Mixed forest", "Shrub/scrub", "Cultivated crops", "Grassland/herbaceous", "Pasture/hay",  "Cultivated crops",  "Cultivated crops",  "Cultivated crops", "Woody wetlands", "Emergent herbaceous wetlands"))

frags.92 <- frags %>%
  filter(year == 1992) %>%
  left_join(classlegend92)
frags.00s <- frags %>%
  filter(year > 1992) %>%
  left_join(classlegend00s)

# Filter out land cover classes of interest only
frags.legend <- bind_rows(frags.92, frags.00s)

frags.filter <- frags.legend %>%
  filter(legend %in% c("Developed, open space", "Developed, low intensity", "Developed, medium intensity", "Developed, high intensity", "Cultivated Crops", "Pasture/hay", "Deciduous forest", "Evergreen forest", "Mixed forest", "Shrub/scrub", "Grassland/herbaceous"))

```

EDA of fragmentation of developed, forested, and grassland land cover in regions of interest

```{r}
# Proportion of landscape deltas
frags.01 <- filter(frags.filter, year == 2001)
frags.11 <- filter(frags.filter, year == 2011)
frags.deltas <- frags.11 %>%
  mutate(delta.PL = prop.landscape - frags.01$prop.landscape) %>%
  mutate(delta.ED = edge.density - frags.01$edge.density) %>%
  mutate(delta.PAR = mean.perim.area.ratio - frags.01$mean.perim.area.ratio)
```

```{r}
theme_set(theme_classic())
# Proportion of landscape
ggplot(filter(frags.filter, year == 2011), aes(x = as.factor(bcr), y = prop.landscape, fill = legend)) + geom_col(position = "stack") + labs(x = "BCR", y = "Proportion of landscape", fill = "Land cover class") + scale_color_viridis_d()
# low proportions include the great lakes
# recalculate proportion as proportion of land cover classes excluding water/maybe ag land

ggplot(frags.deltas, aes(x = as.factor(bcr), y = delta.PL, color = legend)) + geom_point() + geom_hline(yintercept = 0, color = "red") + labs(x = "BCR", y = "Change in proportion of landscape", color = "Land cover class") + scale_color_viridis_d()
# Change relative to starting amount
# Table
# Facet on log/logit scale
```

```{r}
# Edge density
ggplot(filter(frags.filter, year == 2011), aes(x = bcr, y = edge.density, color = legend)) + geom_line() + labs(x = "BCR", y = "Edge Density", color = "Land cover class") + scale_color_viridis_d()
# High edge density indicates high edge length per unit area - highly fragmented

# Edge density deltas
ggplot(frags.deltas, aes(x = as.factor(bcr), y = delta.ED, color = legend)) + geom_line(aes(group = legend)) + geom_hline(yintercept = 0, color = "red") + labs(x = "BCR", y = "Change in edge density", color = "Land cover class") + scale_color_viridis_d()
```

```{r}
# Perimeter area ratio over time
ggplot(filter(frags.filter, year == 2011), aes(x = as.factor(bcr), y = mean.perim.area.ratio, color = legend)) + geom_line(aes(group = legend)) + labs(x = "BCR", y = "Mean P:A", color = "Land cover class") + scale_color_viridis_d()
# High PAR indicates lots of patches - highly fragmented

# Perimeter area ratio deltas
ggplot(frags.deltas, aes(x = as.factor(bcr), y = delta.PAR, color = legend)) + geom_line(aes(group = legend)) + geom_hline(yintercept = 0, color = "red") + labs(x = "BCR", y = "Change in P:A", color = "Land cover class") + scale_color_viridis_d()
```

